Comparing the perfomance impacts of narrow and extended feature spaces for machine learning bots in the game of Schnapsen

### 2018-02-01 Text/paper.tex
title fix
changed in to for
Added 2 pagebreaks for layout
added bibliography and fixed layout of appendix (gone with merging)

\title{Comparing the perfomance impacts of narrow and extended feature spaces in machine learning bots in the game of Schnapsen}
\date{January 25, 2018}
\author{Lucas Faijdherbe, Daan Schrage and Ruben van der Ham\\\\
	\small Intelligent Systems\\
	\small Vrije Universiteit Amsterdam}

@@

\item [ml\_advanced:] 
This bot has a more extended feature space in addition to the \textit{default feature space}. Added are: 
\begin{itemize}
\item Difference between points
\item Difference between pending points
\item Trump card ratio in hand
\item High card ratio in hand (10/A's)
\item Number of trump cards in hand in phase 2
\item Number of cards with same suit as opponents played suit
\item Number of points in hand (squared)
\end{itemize}
This is the bot where we chose to add more of our own thought of features. Point difference was added, since this is a good indicator if a player is leading the game by a high margin. Furthermore, we chose to focus on both trump and high-ranking cards in the hand. The more you have, the better it is. The number of trump cards in phase 2 is also important: If you still got all your trump cards in phase 2 of the game, you have a high chance of beating your opponent. Furthermore the number of cards with same suit as opponents played suit feature was added, alongside the number of points in the player hand squared feature.

@@

\section{References}
1. http://humanparagon.com/strong-and-weak-ai/ \\
2. Lecture 4 of Intelligent Systems, VU, 2018 - \url{https://www.youtube.com/watch?v=nFfjcGKziQE&index=17&list=UUyRRg2ERDkUWKd2p3_KVidA} \\
3. https://towardsdatascience.com/machine-learning-part-3-logistics-regression-9d890928680f \\
4. https://en.wikipedia.org/wiki/Overfitting

\clearpage
\section{Appendix}
\subsection{Overview of machine-learning bots' feature spaces}
\begin{description}
\item[ml\_minimal]

\begin{itemize}

\item Perspective information
\item Player 1's point count
\item Player 2's point count
\end{itemize}
\subsection{ml\_stripped}
\begin{itemize}
\item Perspective information
\item Player 1's point count
\item Player 2's point count
\item The phase
\item The stock size
\item The leader
\item Whose turn it is currently
\item The played card of the opponent
\end{itemize}
\subsection{ml\_enriched}
\begin{itemize}
\item Perspective information
\item Player 1's point count
\item Player 1's point count squared
\item Player 1's point count cubed
\item Player 2's point count
\item Player 2's point count squared
\item Player 2's point count cubed
\item Player 1's pending points
\item Player 1's pending points squared
\item Player 1's pending points cubed
\item Player 2's pending points
\item Player 2's pending points squared
\item Player 2's pending points cubed
\item The trump suit
\item The phase
\item The stock size
\item The leader
\item Whose turn it is currently
\item The played card of the opponent
\end{itemize}
\subsection{ml\_advanced}
\begin{itemize}
\item Perspective information
\item Player 1's point count
\item Player 2's point count
\item Player 1's pending points
\item Player 2's pending points
\item Difference between player 1 and player 2's point count
\item Difference between player 1 and player 2's pending point count
\item The trump suit
\item Trump card ratio (Number of trump cards in hand / total number of cards in hand * 100)
\item High-ranking card ratio (Number of 10's and A's in hand / total number of cards in hand * 100)
\item The phase
\item Number of trump cards in hand in phase 2
\item The stock size
\item The leader
\item Whose turn it is currently
\item The played card of the opponent
\item Number of cards in same suit as opponent's played suit in hand.
\item Number of points in hand squared ($card1^2$ + $card2^2$ ...)
\end{itemize}

\subsection{ml\_combined}
\begin{itemize}
\item Perspective information
\item Player 1's point count
\item Player 1's point count squared
\item Player 1's point count cubed
\item Player 2's point count
\item Player 2's point count squared
\item Player 2's point count cubed
\item Player 1's pending points
\item Player 1's pending points squared
\item Player 1's pending points cubed
\item Player 2's pending points
\item Player 2's pending points squared
\item Player 2's pending points cubed
\item Difference between player 1 and player 2's point count
\item Difference between player 1 and player 2's pending point count
\item The trump suit
\item Trump card ratio (Number of trump cards in hand / total number of cards in hand * 100)
\item High-ranking card ratio (Number of 10's and A's in hand / total number of cards in hand * 100)
\item The phase
\item Number of trump cards in hand in phase 2
\item The stock size
\item The leader
\item Whose turn it is currently
\item The played card of the opponent
\item Number of cards in same suit as opponent's played suit in hand.
\item Number of points in hand squared ($card1^2$ + $card2^2$ ...)
\end{itemize}
\end{description}
\section{Worksheet I}
\begin{enumerate}

@@

\end{enumerate}
\section{Worksheet II}
\begin{enumerate}
\item The added clause that makes the knowledge base unsatisfiable:
\begin{lstlisting}[language=bash]
kb.add_clause(~B, ~C)
\end{lstlisting}

@@

\item 
\end{enumerate}
\section{Worksheet III}
\begin{enumerate}
\item The added code:
\begin{lstlisting}[language=bash]
value = self.heuristic(next_state)

>>>>

\title{Comparing the perfomance impacts of narrow and extended feature spaces for machine learning bots in the game of Schnapsen}
\date{January 25, 2018}
\author{Lucas Faijdherbe, Daan Schrage and Ruben van der Ham\\\\
	\small Intelligent Systems\\
	\small Vrije Universiteit Amsterdam}

@@

\clearpage
\item [ml\_advanced:] 
This bot has a more extended feature space in addition to the \textit{default feature space}. Added are: 
\begin{itemize}
\item Difference between points
\item Difference between pending points
\item Trump card ratio in hand
\item High card ratio in hand (10/A's)
\item Number of trump cards in hand in phase 2
\item Number of cards with same suit as opponents played suit
\item Number of points in hand (squared)
\end{itemize}
This is the bot where we chose to add more of our own thought of features. Point difference was added, since this is a good indicator if a player is leading the game by a high margin. Furthermore, we chose to focus on both trump and high-ranking cards in the hand. The more you have, the better it is. The number of trump cards in phase 2 is also important: If you still got all your trump cards in phase 2 of the game, you have a high chance of beating your opponent. Furthermore the number of cards with same suit as opponents played suit feature was added, alongside the number of points in the player hand squared feature.

@@

\section{References}

\begin{enumerate}
\item What's The Difference Between Strong and Weak AI?, \url{http://humanparagon.com/strong-and-weak-ai/}.

\item Schlobach, S.: Intelligent Systems 2018 L4P2 More on Informed Search, \url{https://www.youtube.com/watch?v=nFfjcGKziQE}.

\item Chandrayan, P.: Machine Learning Part 3 : Logistic Regression – Towards Data Science, \url{https://towardsdatascience.com/machine-learning-part-3-logistics-regression-9d890928680f}.

\item Overfitting, \url{https://en.wikipedia.org/wiki/Overfitting}.\\

\end{enumerate}

\clearpage
\section{Appendix}
\subsection{Overview of machine-learning bots' feature spaces}
\begin{description}
\item[ml\_minimal] \hfill
\begin{itemize}
\item Perspective information
\item Player 1's point count
\item Player 2's point count
\end{itemize}
\item[ml\_stripped] \hfill
\begin{itemize}
\item Perspective information
\item Player 1's point count
\item Player 2's point count
\item The phase
\item The stock size
\item The leader
\item Whose turn it is currently
\item The played card of the opponent
\end{itemize}
\item[ml\_enriched] \hfill
\begin{itemize}
\item Perspective information
\item Player 1's point count
\item Player 1's point count squared
\item Player 1's point count cubed
\item Player 2's point count
\item Player 2's point count squared
\item Player 2's point count cubed
\item Player 1's pending points
\item Player 1's pending points squared
\item Player 1's pending points cubed
\item Player 2's pending points
\item Player 2's pending points squared
\item Player 2's pending points cubed
\item The trump suit
\item The phase
\item The stock size
\item The leader
\item Whose turn it is currently
\item The played card of the opponent
\end{itemize}
\item[ml\_advanced] \hfill
\begin{itemize}
\item Perspective information
\item Player 1's point count
\item Player 2's point count
\item Player 1's pending points
\item Player 2's pending points
\item Difference between player 1 and player 2's point count
\item Difference between player 1 and player 2's pending point count
\item The trump suit
\item Trump card ratio (Number of trump cards in hand / total number of cards in hand * 100)
\item High-ranking card ratio (Number of 10's and A's in hand / total number of cards in hand * 100)
\item The phase
\item Number of trump cards in hand in phase 2
\item The stock size
\item The leader
\item Whose turn it is currently
\item The played card of the opponent
\item Number of cards in same suit as opponent's played suit in hand.
\item Number of points in hand squared ($card1^2$ + $card2^2$ ...)
\end{itemize}

\item[ml\_combined] \hfill
\begin{itemize}
\item Perspective information
\item Player 1's point count
\item Player 1's point count squared
\item Player 1's point count cubed
\item Player 2's point count
\item Player 2's point count squared
\item Player 2's point count cubed
\item Player 1's pending points
\item Player 1's pending points squared
\item Player 1's pending points cubed
\item Player 2's pending points
\item Player 2's pending points squared
\item Player 2's pending points cubed
\item Difference between player 1 and player 2's point count
\item Difference between player 1 and player 2's pending point count
\item The trump suit
\item Trump card ratio (Number of trump cards in hand / total number of cards in hand * 100)
\item High-ranking card ratio (Number of 10's and A's in hand / total number of cards in hand * 100)
\item The phase
\item Number of trump cards in hand in phase 2
\item The stock size
\item The leader
\item Whose turn it is currently
\item The played card of the opponent
\item Number of cards in same suit as opponent's played suit in hand.
\item Number of points in hand squared ($card1^2$ + $card2^2$ ...)
\end{itemize}
\end{description}
\subsection{Worksheet I}
\begin{enumerate}

@@

\end{enumerate}
\subsection{Worksheet II}
\begin{enumerate}
\item The added clause that makes the knowledge base unsatisfiable:
\begin{lstlisting}[language=bash]
kb.add_clause(~B, ~C)
\end{lstlisting}

@@

\item 
\end{enumerate}

\clearpage
\subsection{Worksheet III}
\begin{enumerate}
\item The added code:
\begin{lstlisting}[language=bash]
value = self.heuristic(next_state)
